---
title: "PHYSICAL ACTIVITY ENVIRONMENT AND PERFEC/TENACI"
author: "Serrano, M. F."
date: "Today"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
		      #results = "hide", 
		      fig.height = 8.5,
		      fig.width = 8.5,
			message = FALSE)
## LOAD ACTIVITY REGRESSION ANALYSES SCRIPT
if(!exists('res_activity_diff')) source('~/p/oros/regres_activi.R')
```

# INTRODUCTION.
Previous analyses showed that enviromental variables, such as school, have significant effects on IQ and Perf. variables.

For this reason:

- we explore the effect of the enviroment variables that promote physiscal activity in schools on IQ and Perf variables.
- Variables related to how the evironment of schools devoted to physical activity were collected.
- They are named and defined in the following section.

# DOCUMENTATION

## VARIABLES DE AMBIENTE PARA ACTIVIDAD FISICA
El sufijo aaf_ en todas las variables abrevia:
Ambiente (para) Actividad Fisica.

- aaf_

### OBSERVED VARIABLES
| Nombre d Variable | Definicion | Valores |
| ----------------- | ---------- | ------- |
| aaf_t_pe_class| t Tiempo en clase de ed. fisica	| Minutos por semana |
| aaf_t_recess	| t Tiempo de recreo                    | Minutos por semana |
| aaf_population_size | Poblacion total de la escuela	| 1 - 4 = chica, mediana, grande, muy_gr |
| aaf_s_size	| s Tamannho d Espacio para actividad	| 0 - 4 = no_hay, chico, mediano, grande, muy_gr|
| aaf_s_avail	| s Espacio esta disponible o no        | 0, 1 |
| aaf_s_used	| s Espacio se usa o no                 | 0, 1 |
| aaf_s_shape	| s Forma del espacio                   | rectang, triang, irregular, other |

### COMPUTED VARIABLES
| Computed Variable | Definition | Formula |
| ----------------- | -----------| ------- |
| aaf_t_sum_total | total aggregated class+recess time  | t_sum = t_class + t_recess    |
| aaf_ratio_s_pop | space-size population-size ratio    | ratio_sp = s_size / pop_size     |
| aaf_indica_rec_t_s_p | time_space/population Indicator: Product of recess-time times the space/population ratio   | indica_tsp = t_recesss * ratio_sp          |
| aaf_indica_sum_t_s_p | total time_space/population Indicator: Product of aggregated time (class+recess) times the space/population ratio   | indica_sum_tsp = t_sum * ratio_sp  |


# ANALYSES.
## LINEAR REGRESSSION ASSOCIATION ANALYSES.
### PHYSICAL ACTIVITY ENVIROMENT VS IQ/FROST/OROS VARIABLES.
Six significant associations were found.

- `r length(res_activity_corr[['pr2']])` significant differences were found.
- The strongest association (r>0.6) was found between IQ and physical activity aggregated time.
- Of particular interest:
- Association between Oros's PSP social dimension and aggregated time was found.

The following is the list of significant results, listing the name of variables, pvalue and Adjusted Rsquared.
```{r show_corr_function, include = TRUE}
show_pars(res_activity_corr, decreasing = FALSE)
```
```{r draw_grid_corr, echo=FALSE, results='hide', fig.keep='all', fig.width = 6.5, fig.height = 6.5}
grid.draw(res_activity_corr[['grid']])
```

### DIFFERENCES BETWEEN GROUPS ANALYSES.
By using the physical activity as groupping variables:

- `r actdiffsiglength` significant differences were found.
- Most of them with small effects.
- However, the associtiaion between IQ and Oros's  PSP social dimension was replicated.

The following is the list of significant results, listing the name of variables, pvalue and Adjusted Rsquared.
```{r show_diff_function, include = TRUE}
## PRINT THE PARAMETERS WITH THE SHOW_PARS FUNTION
show_pars(res_activity_diff, decreasing = FALSE)
```
```{r draw_grid_diff, echo=FALSE, results='hide', fig.keep='all', fig.width = 10.5, fig.height = 10.5}
## DRAW THE DIFFERENCES BY CATEGORICAL VARIABLES GRID
grid.draw(res_activity_diff[['grid']])
```

## VARIABLE OF INTEREST, OROS'S PSP SOCIAL DIMENSION, GRAPHICS IN DETAIL.
### PSP-TIME ASSOCIATION.

```{r corr_sumt_psp, include = TRUE, fig.width = 6.5, fig.height = 6.5}
## SHOW THE PSP CORRELATION GRAPH
res_activity_corr[['plots']][['PSP_Total_~_aaf_t_sum_total']]
```

### DIFFERENCES IN PSP SOCIAL DIMENSION.

```{r diff_sumt_psp, include = TRUE, fig.width = 6.5, fig.height = 6.5}
## SHOW THE PSP CORRELATION GRAPH
res_activity_diff[['plots']][['PSP_Total_~_aaf_t_sum_total']]
```
